---
title: "Training Model For Predicting How Well An Individual Perform A Particular Exercise"
output: html_document
---

## Introduction
This model is meant to quantify how well an individual perform for a particular activity. Data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants are used. The participants were asked to perform barbell lifts correctly and incorrectly in 5 different ways.

More information is available from the website here: http://groupware.les.inf.puc-rio.br/har
(see the section on the Weight Lifting Exercise Dataset).

The training data for this project are from here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data used are from here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

For reproducibility, seed values are set and for prediction, random forest is used for its accuracy although it is more difficult to interpret.

## Loading Libraries
```{r cache=TRUE, echo=TRUE, eval=FALSE}
rm(list=ls())
library(caret)
library(kernlab)
library(AppliedPredictiveModeling)
library(randomForest)
library(ggplot2)
library(Metrics)
```

## Data Preparation
Loading the data:
```{r cache=TRUE, echo=TRUE, eval=FALSE}
setwd("D:/R/pml/")
train_url<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
test_url<-"https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(train_url,destfile="~/training.csv")
download.file(test_url,destfile="~/testing.csv")
train<-read.csv("~/training.csv", na.strings=c("NA","#DIV/0!",""))
test<-read.csv("~/testing.csv",na.strings=c("NA","#DIV/0!",""))
```

Preparing and cleaning the data:

1) Subsetting required data for the analysis. Removing train data column 1 to 7, user name, window, and time. NAs are also removed for a more accurate prediction.

2) Using the function createDataPartition to split the training set into training_train and training _test sets, creating a single 75% and 25% split of the data. If the y argument to this function is a factor, the random sampling occurs within each class and should preserve the overall class distribution of the data.
```{r cache=TRUE, echo=TRUE, eval=FALSE}
training<-train[,-c(1:7)]
set.seed(100)
inTrain<-createDataPartition(training$classe, p=0.75, list=FALSE)
training_train<-training[inTrain,]
training_test<-training[-inTrain,]
dim(training_train)
training_train<-as.data.frame(training_train)
clean_train<-training_train[,colSums(is.na(training_train))==0]
clean_test<-training_test[,colSums(is.na(training_test))==0]

```

## Build Model & Analysis

```{r cache=TRUE, echo=TRUE, eval=FALSE}
set.seed(150)
tr<-trainControl(method="cv", number=5)
#first prediction model
modelFit1 <- train(classe ~.,data=clean_train, method="rf", trControl=tr)
prediction1<-predict(modelFit1, clean_train)
print(modelFit1)
summary(modelFit1$finalModel)
confusionMatrix(predict(modelFit1, newdata=clean_test), clean_test$classe)
table(prediction1, clean_train$classe)
```

## Prediction
```{r cache=TRUE, echo=TRUE, eval=FALSE}
#predict train_test data
modelFit2<-train(classe~., data=clean_test, method="rf", trControl=tr)
prediction2<-predict(modelFit1, newdata=clean_test)
table(prediction2, clean_test$classe)
sum(clean_test$classe==predict(modelFit1, clean_test)) / nrow(clean_test)
test<-test[,colSums(is.na(test))==0]
test<-test[,-c(1:7)]
#predict test data
prediction3<-predict(modelFit2, newdata=test)

```


## Generate Problem ID Files
```{r cache=TRUE, echo=TRUE, eval=FALSE}
#coursera code 
pml_write_files = function(x){
        n = length(x)
        for(i in 1:n){
                filename = paste0("problem_id_",i,".txt")
                write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
        }
}

pml_write_files(prediction3)


```

## Conclusion
Out of sample error is 0.993. Confusion matrix is relatively accurate.

## Appendix
Figure 1: Plot of modelFit
```{r cache=TRUE, echo=TRUE, eval=FALSE}
ggplot(modelFit1)
```

Figure 2: Importance of the variables
```{r cache=TRUE, echo=TRUE, eval=FALSE}

resize.win <- function(Width=6, Height=6)
{
        # windows
        dev.off(); 
        windows(record=TRUE, width=Width, height=Height)
}
resize.win(5,5)
plot(rnorm(100))
resize.win(10,10)
plot(rnorm(100))
plot(varImp(modelFit1))
```

Figure 3: Prediction plot
```{r cache=TRUE, echo=TRUE, eval=FALSE}
qplot(predict(modelFit1,clean_test), classe, data=clean_test)
```